---
title: "CIRCULATE analysis"
output: html_notebook
---
```{r}
library(dplyr)
library(survival)
library(survminer)
```

# DFS analysis

## Ext.Fig4:pre-op all stages
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/pre-op"))
cir_data<-read.csv("pre-op.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Fig2A:post-op all stages
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/post-op"))
cir_data<-read.csv("post-op.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig3A:post-op stage I

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/post-op"))
cir_data<-read.csv("post-opstageI.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig3B:post-op stage II

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/post-op"))
cir_data<-read.csv("post-opstageII.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig3C:post-op stage III

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/post-op"))
cir_data<-read.csv("post-opstageIII.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig3D:post-op stage IV

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/DFS analysis/post-op"))
cir_data<-read.csv("post-opstageIV.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

# Fig2B:Multivariate analysis for recurrence all stage II-III
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Multivariate analysis"))
cir_data<-read.csv("MVAII-III.csv")
cir_data$sex <- as.factor(cir_data$sex)
cir_data$age1 <- as.factor(cir_data$age1)
cir_data$PS <- as.factor(cir_data$PS)
cir_data$pT <- as.factor(cir_data$pT)
cir_data$pN <- as.factor(cir_data$pN)
cir_data$MSS <- as.factor(cir_data$MSS)
cir_data$BRAF <- as.factor(cir_data$BRAF)
cir_data$RAS <- as.factor(cir_data$RAS)
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
cox_fit <- coxph(surv_object ~ ctDNA + sex + age1 + PS + pT + pN + MSS + BRAF + RAS, data=cir_data)
ggforest(cox_fit,data = cir_data)
test.ph <- cox.zph(cox_fit)
test.ph
```

# Ext.Fig5D:Multivariate analysis for recurrence ctDNApos 4 weeks stage II-IV
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Multivariate analysis"))
cir_data<-read.csv("MVA2ACT.csv")
cir_data$sex <- as.factor(cir_data$sex)
cir_data$age1 <- as.factor(cir_data$age1)
cir_data$PS <- as.factor(cir_data$PS)
cir_data$Stage <- as.factor(cir_data$Stage)
cir_data$MSS <- as.factor(cir_data$MSS)
cir_data$BRAF <- as.factor(cir_data$BRAF)
cir_data$RAS <- as.factor(cir_data$RAS)
cir_data$ACT <- as.factor(cir_data$ACT)
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cir_data$age1 = relevel(cir_data$age1, ref = "1")
cir_data$PS = relevel(cir_data$PS, ref = "1")
cir_data$MSS = relevel(cir_data$MSS, ref = "MSS")
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
cox_fit <- coxph(surv_object ~ ACT + sex + age1 + PS + Stage + MSS + BRAF + RAS, data=cir_data)
ggforest(cox_fit,data = cir_data)
test.ph <- cox.zph(cox_fit)
test.ph
```

# Fig4:4-12 wks Dynamics analysis
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Dynamics analysis"))
wks_data<-read.csv("4-12wks.csv")
surv_object_wrs<-Surv(time = wks_data$DFS, event = wks_data$Event)
KM_curve_wks <- survfit(surv_object_wrs ~ strata(ctDNA_dynamics), data = wks_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve_wks, data = wks_data, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve_wks, times= c(0,6,12,18,21,24))
myplot<-ggsurvplot(KM_curve_wks, data = wks_data, conf.int = TRUE,risk.table = TRUE,break.time.by=3,xlim = c(3,24))
myplot$plot <- myplot$plot + scale_x_continuous(breaks = c(3,9,15),labels = c(3,9,15))
myplot
```

## 4-12 wks posnegtonegpos
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Dynamics analysis"))
cir_data<-read.csv("pntonp.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA_dynamics, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA_dynamics, data=cir_data)
cir_data$ctDNA_dynamics = relevel(cir_data$ctDNA_dynamics, ref = "Positive>Negative")
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## 4-12 wks posnegtonegneg
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Dynamics analysis"))
cir_data<-read.csv("pntonn.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA_dynamics, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA_dynamics, data=cir_data)

ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## 4-12 wks posnegtopospos
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Dynamics analysis"))
cir_data<-read.csv("pntopp.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA_dynamics, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA_dynamics, data=cir_data)
cir_data$ctDNA_dynamics = relevel(cir_data$ctDNA_dynamics, ref = "Positive>Negative")
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## 4-12 wks negnegtonegpos

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Dynamics analysis"))
cir_data<-read.csv("nntonp.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ctDNA_dynamics, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cox_fit <- coxph(surv_object ~ ctDNA_dynamics, data=cir_data)

ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```


# 2m filter:DFS post-op ctDNA pos ACT vs noACT 

## Ext.Fig6B:high-risk stage II and III ctDNA neg
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("HighriskII-IIIneg(1).csv")
cir_data$DFS=cir_data$DFS-2
cir_data <- cir_data[cir_data$DFS>=0,]
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cir_data$MSS <- as.factor(cir_data$MSS)
cir_data$pStage <- as.factor(cir_data$pStage)
cir_data$age1 <- as.factor(cir_data$age1)
cox_fit <- coxph(surv_object ~ ACT + PS + age1 + MSS + pStage, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig6A:high-risk stage II and III ctDNA pos
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("HighII-IIIMRDposnew.csv")
cir_data$DFS=cir_data$DFS-2
cir_data <- cir_data[cir_data$DFS>=0,]
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cox_fit <- coxph(surv_object ~ ACT + PS + Gender, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```


# DFS post-op ctDNA pos ACT vs noACT (no filter)
## Ext.Fig5A:highrisk stage II
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("highriskIIpos.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=3,xlim = c(0,24))
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cox_fit <- coxph(formula=surv_object~ ACT + PS + Gender, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```


## Ext.Fig5B:stage III
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("stageIIIpos.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=3,xlim = c(0,24))
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cox_fit <- coxph(surv_object ~ ACT + PS + Gender, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig5C:stage IV
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("stageIVpos.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=3,xlim = c(0,24))
summary(KM_curve, times= c(0,6,12,18, 21, 24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cox_fit <- coxph(surv_object ~ ACT + PS + Gender, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```


## Fig3B:high-risk stage II and III ctDNA neg
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("HighriskII-IIIneg(1).csv")
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cir_data$MSS <- as.factor(cir_data$MSS)
cir_data$pStage <- as.factor(cir_data$pStage)
cir_data$age1 <- as.factor(cir_data$age1)
cox_fit <- coxph(surv_object ~ ACT + PS + age1 + MSS + pStage, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Fig3A:high-risk stage II and III ctDNA pos
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/ACT analysis"))
cir_data<-read.csv("HighII-IIIMRDposnew.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cox_fit <- coxph(surv_object ~ ACT + PS + Gender, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

# Fig5B:MRD pos cleared vs not cleared on ACT

```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/Clearance data"))
cir_data<-read.csv("MRDposclearedvsnocleared.csv")
surv_object<-Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ Cleared, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$Cleared = relevel(cir_data$Cleared, ref = "Yes")
cox_fit <- coxph(surv_object ~ Cleared, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

# StageIV DFS post-op NAC treatment
## Ext.Fig7A:NAC treated MRD pos ACT vs noACT (StageIV)-2m filter
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/NAC analysis"))
cir_data<-read.csv("stageIVposNAC.csv")
cir_data$DFS=cir_data$DFS-2
cir_data <- cir_data[cir_data$DFS>=0,]
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cir_data$MSS <- as.factor(cir_data$MSS)
cir_data$age1 <- as.factor(cir_data$age1)
cir_data$PS <- as.factor(cir_data$PS)
cir_data$sex <- as.factor(cir_data$sex)
cox_fit <- coxph(surv_object ~ ACT + sex, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

## Ext.Fig7B:no NAC treated MRD pos ACT vs noACT (StageIV)-2m filter
```{r}
setwd (("/Users/ssharma/Downloads/Circulate analysis/Updated data/NAC analysis"))
cir_data<-read.csv("stageIVposnoNAC.csv")
cir_data$DFS=cir_data$DFS-2
cir_data <- cir_data[cir_data$DFS>=0,]
surv_object <- Surv(time = cir_data$DFS, event = cir_data$Event)
KM_curve <- survfit(surv_object ~ ACT, data = cir_data, conf.int=0.95,conf.type="log-log")
ggsurvplot(KM_curve, data = cir_data, pval = TRUE, conf.int = TRUE,risk.table = TRUE,break.time.by=6)
summary(KM_curve, times= c(0,6,12,18,21,24))
cir_data$ACT = relevel(cir_data$ACT, ref = "YES")
cir_data$age1 <- as.factor(cir_data$age1)
cir_data$PS <- as.factor(cir_data$PS)
cir_data$sex <- as.factor(cir_data$sex)
cox_fit <- coxph(surv_object ~ ACT + sex + PS, data=cir_data)
ggforest(cox_fit,data = cir_data)
summary(cox_fit)
```

# Extended Figure 2
```{r}

require(openxlsx)
require(data.table)
require(ggplot2)

dt.timepoints <- read.xlsx('NMED sup table.xlsx', sheet = 'Timepoints')
setDT(dt.timepoints)

dt.pos_baseline <- dt.timepoints[VISITTYPE=='Initial Collection' & Result=='POSITIVE',]
dt.pos_4wMRD    <- dt.timepoints[VISITTYPE=='Post-Surgery' & Timepoint=='4WEEK' & Result=='POSITIVE',]

p.ExtFig2A <-
  dt.pos_baseline %>%
  ggplot(aes(x=Stage,y=MTM,fill=Stage)) +
  scale_y_log10()+
  geom_violin(alpha=0.2)+
  leg.no+ylab('MTM/mL')+
  geom_boxplot(alpha=0.6)


t.test(dt.pos_baseline[Stage=='I',  log10(MTM)],dt.pos_baseline[Stage=='II',log10(MTM)])
t.test(dt.pos_baseline[Stage=='II', log10(MTM)],dt.pos_baseline[Stage=='III',log10(MTM)])
t.test(dt.pos_baseline[Stage=='III',log10(MTM)],dt.pos_baseline[Stage=='IV',log10(MTM)])


p.ExtFig2B <-
  dt.pos_4wMRD %>%
  ggplot(aes(x=Stage,y=MTM,fill=Stage)) +
  scale_y_log10()+
  leg.no+xlab('Stage')+ylab('MTM/mL')+
  geom_violin(alpha=0.2)+
  geom_boxplot(alpha=0.6)


p.ExtFig2C <-
  dt.pos_baseline %>%
  ggplot(aes(x=StageMRD,y=MTM,fill=Stage)) +
  scale_y_log10()+
  geom_violin(alpha=0.2)+
  leg.no+xlab('Stage and MRD status')+ylab('MTM/mL')+
  geom_boxplot(alpha=0.6)


p.ExtFig2D <-
  dt.pos_4wMRD[Stage!='I',] %>%
  ggplot(aes(x=StageClr,y=MTM,fill=Stage)) +
  scale_y_log10(labels = scales::comma)+
  leg.no+xlab('Stage and Cleared status')+ylab('MTM/mL')+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())+
  geom_violin(alpha=0.2)+
  geom_boxplot(alpha=0.6)

t.test(dt.pos_4wMRD[StageClr=='II Yes', log10(MTM)],dt.pos_baseline[StageClr=='II No', log10(MTM)])
t.test(dt.pos_4wMRD[StageClr=='III Yes',log10(MTM)],dt.pos_baseline[StageClr=='III No',log10(MTM)])
t.test(dt.pos_4wMRD[StageClr=='IV Yes', log10(MTM)],dt.pos_baseline[StageClr=='IV No', log10(MTM)])

```

